---
title: "R OOP"
author: "Chao-Jen Wong"
date: "January 23, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Terminalogy
Method
Dispatch

S3 - linear hierarch


## Create class

```{r passenger}
#' superclass
setClass("Passenger", 
  representation(fist_name = "character",
                 last_name = "character",
                 origin = "character",
                 destination = "character")
#' subclass inheriting properties of superclass
setClass("FreqFlyer",
  representation(contains = "Passenger"),
                 ffnumber = "numeric")

#getClass("FreqFlyer")   
#subClassNames("Passenger")
#superClassNames("FreyFlyer")              
```

## S3  - dispatch
Method - specialized function applied to instances of one or more classes.
Dispatch - process of determining the appropriate method to invoke 

S3  
- very little code is needee  
- only examines the first argument and dispatches depending on its class
- dispatch by function naming convention that is describred .*  


```{r show_passenger}
show.Passenger <- function(x) {
  NextMethod("show", x)
}
```

## Self-describing data
A self-describing data object can be used to carry out a reasonable analysis of the data.