---
title: "R OOP"
author: "Chao-Jen Wong"
date: "January 23, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Terminalogy
Object - data object which has a class attribute
heritance
polymorphism
Dispatch
Generic function
Method

R OOP  
- S3: generic functions, linear hierarch 
- S4: classes and generic functions, polymorphism  

## S3 system
- dispatch system for generic functions 
- loessly define classes
- easy to describe and use
- less code is needee  
- only examines the first argument and dispatches depending on its class
- dispatch by function naming convention that is describred .*  

Start with an glm class and pring (generic function) and method print.glm
```{r S3_method}
print
```
A singel expression which is the call to `UseMethod`, which is the mechanism that 
helps to dispatch to the appropriate method.
`print()` is a generic function. A generic function does dispatch throughed a call to `
UseMethod` that initaites the dispatch ona single argument, usually the first one.


```{r example_glm, eval=FALSE}
methods(print)
```

```
 [90] print.ftable*                                     
 [91] print.function                                    
 [92] print.getAnywhere*                                
 [93] print.glm*                                        
 [94] print.hclust*                                  
```

```{r exp_glm}
#getS3method("print", "glm")
#methods(class="glm")
```

## S3: dummy example
Lossely defien an S3 class `Passenger`:

```{r s3_classa}
Lossely define an S3 class
# Define S3 class
x <- list(name="John Alber", origin="SEA", destination="KCM")
class(x) <- "Passenger"
```

## S3: dummy example
Generic functions and methods

```{r s3_method}
# define foo generic funciton
foo <- function(x, ...) UseMethod("foo")
# define default
foo.default <- function(x, ...) {
   print(x)
}
# define method applied to Passenger
foo.Passenger <- function(x) {
  cat(paste0("Name: ", x$name, "\n"))
  cat(paste0("Origin: ", x$origin, "\n"))
  cat(paste0("Dest: ", x$destination), "\n")
  NextMethod()
}
# getS3method("foo", "Passenger")
```

`NextMethod()` transfers contorl to the next most specific method, which is `foo.default()`.

When `NextMethod("print")` applied to an object with class attribute `Passenger`, the system 
search for a function called `print.Passenger`. If it finds it, applie to the object. Otherwise,
it uses `print.default()`, if it exists.

## S4 system

- overcome the deficientices of S3 system  
  - multiple dispatch  
  - object validation  
  - coercion system
  - explicit representation of classes virtual/concrete classes
  - multiple inheritance
  - class unions



## S4: class
Goal: define self-describing data object can be used to carry out a reasonable 
analysis of the data

Concrete Class
- `class` 
- `presentation`  
- `contains`    
- `prototype`  
- `validity`

## S4: concrete class
Establish two classes: `Passenger` and `FreqFlyer`

```{r define_passenger}
#' class and representation (slots)
setClass("Passenger", 
  slots = c(name = "character",
            origin = "character",
            destination = "character"),
  prototype = prototype(name = NA_character_,
                        origin = NA_character_,
                        destination = NA_character_))                        
```

```{r subclass}
#' contains
#' subclass inheriting properties of superclass
setClass("FreqFlyer",
  contains = "Passenger",
  slots= c(ffnumber = "numeric"))
```

## S4: classes
coercion, and accessories.

```{r example_passenger}
#' S4 classes
x <- new("Passenger", name="Josephine Andre", 
         origin = "SEA",
         destination = "KCM")
x
# inherits(x, "Passenger")    
# showClass("Passenger")
# slotNames("Passenger")      
```

## S4: Validation
Validation (`setValidity()`): each slot has length 1.

```{r passenger_validate}

```

## S4: generic functions/methods
constructor, validation, initialization, getters, setters, show, 
coercion, and accessories.

### Define `show` method
```{r passenger_show}
#` methods("show")
#` show is already a generic functions for S4 methods 
setMethod("show", signature="Passenger",
          function(object) {
          cat(paste0("name: ", object@name, "\n"))
          cat(paste0("origin: ", object@origin, "\n"))
          cat(paste0("destination: ", object@destination, "\n"))
          })
show(x)
# showMethods("show")
# selectMethod("show", "Passenger")
```

## S4: constructor
Instead of `new()`:

```{r constructor_passenger}
#` constructor
Passenger <- function(name, origin, destination) {
    new("Passenger", name=name, origin=origin, destination=destination)
}
x <- Passenger(name="Josie Andre", origin="SEA", destination="NYC")
```

## S4: accessors
- prefer user to access the internal of an object via accessors (getters and setters). R does not have provision of encapsulation.
- crafted setters ensure the object remains valid 
- allow developers to make object more memory compact and efficient

```{r getter_passenger}
#' getter
setGeneric("name", function(object) standardGeneric("name"))
setMethod("name", signature("Passenger"),
    function(object) {object@name})
#' setter (replacement methods)
setGeneric("name<-", function(object, value) standardGeneric("name<-"))
setMethod("name<-", signature("Passenger"),
    function(object, value) {
        # validate value
        if (is.character(value))
            object@name <- value
        if (!is.character(value))
            warning("`name` in a Passenger class must be character")
        object
    })

name(x) <- "Josie W. Andre"
name(x) <- c("A", "B")
```

## S4: coercion
Coerce `Passenger` to `data.frame`

```{r coerce_passenger}
setAs("Passenger", "data.frame", 
      function(from)
          data.frame(name = from@name,
                     origin = from@origin,
                     destination = from@destination))
as(x, "data.frame")                     
```

## S4: virtual classes 
Cannot be instantiated. Used as the parent of one or more concrete class. 

```{r virtual_class, eval=FALSE}
setClass("A", representation("VIRTUAL")) # can have slots or not
# can define generic functions delicate to the virtual class
setClass("A1", contains = "A", slots = ...add slots here...)
setClass("A2", contains = "A", slots = ...add slots here...)
```

## S4: virtual class example
`GenomicRanges`, superclass of widely-used subclassed in BioConductor

```{r example_vartual}
suppressPackageStartupMessages(library(GenomicRanges))
showClass("GenomicRanges")
```

- Subclasses such as `GRanges` and `GPos` inherite the properties of the superclass.
- Methods are inherited and can be applied to `GRanges` adn `GPos`

```{r name_method, eval=FALSE}
selectMethod("names", "GRanges")
```

## S4: multiple inheritance
Powerful but code with caution.

```{r compressed_iranges}
showClass("COmpressedIRangesList")
```

## S4: class unions
## Reference classes
pass-by-copy and pass-by-reference semantic

## Applications - Bioconductor's GRanges/SummarizedExperiments

## Install Bioconductor package

```{r biocmanager, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("S4Vectors", version = "3.8")
```

## Reference

- The _R Programming fro Bioinformatics_ by Robert Gentleman  
- The _Writing R Extensions manual_ 
- A quick overview of S4 class system by Herv´e Pag`es,
  https://bioconductor.org/packages/release/bioc/vignettes/S4Vectors/inst/doc/S4QuickOverview.pdf    
- Bioconductor training courses:
  https://master.bioconductor.org/help/course-materials/2017/Zurich/S4-classes-and-methods.html